Adopting agile in the large: motivations, success factors and challenges

### 2014-05-18 journal/article.tex
Reasons to change refactored

As stated by Research Question 1 we investigated the reasons organizations start
a transformation process towards agile software development. Most primary
studies mentioned reasons for starting the transformation, and only 7 cases of
42 did not mention any reasons. In the cases giving no specific reason to change
it was simply mentioned that management had decided to change the organization.
The identified reasons for organizations to change are presented in table
\ref{table:reasonstochange}.

@@

\begin{table}
    \centering
    \begin{tabular}{ p{1.2cm} p{1.2cm} l }
        \toprule
        \multicolumn{3}{l}{Reasons to change}  \\
        \midrule

        \multicolumn{3}{l}{External issues: Need to respond to changes in business}  \\

        &  \multicolumn{2}{l}{Demand for more frequent deliveries} \\
        &  &  Time To Market must lessen  \\
        &  &  Long cycle times are seen as problematic  \\

        &  \multicolumn{2}{l}{Something must be done in order to remain competitive} \\
        &  &  Change is required to stay in business  \\

        &  \multicolumn{2}{l}{Customers demand improvement} \\

        &  \multicolumn{2}{l}{Demand for flexibility or responsiveness} \\
        &  &  Demand to accomodate changes efficiently  \\
        &  &  Better responsiveness to customer requests  \\

        \multicolumn{3}{l}{\rule{0pt}{4ex} Internal issues: Problems in managing software development}  \\

        &  \multicolumn{2}{l}{General problems with management of software development} \\
        &  &  Management dysfunction  \\
        &  &  Problems with motivation and morale  \\
        &  &  Problems coordinating a growing organization  \\

        &  \multicolumn{2}{l}{Problems with overhead and excess in process}  \\
        &  &  Excess, slowness and costs caused by process  \\
        &  &  The current process is being ignored  \\

        &  \multicolumn{2}{l}{Problems with planning or keeping schedules}  \\

        &  \multicolumn{2}{l}{Too long cycle time before feedback}  \\

        &  \multicolumn{2}{l}{Demand to improve product quality}  \\

        \multicolumn{3}{l}{\rule{0pt}{4ex} Communication and collaboration issues}  \\

        &  \multicolumn{2}{l}{Communication and collaboration dificulties}  \\
        &  &  Problems in communication across and within projects  \\
        &  &  Teams work in silos or isolation  \\

        &  \multicolumn{2}{l}{Requirements engineering issues}  \\
        &  &  Requirements are not understood  \\
        &  &  Unawareness of what is important  \\

        &  \multicolumn{2}{l}{Demand of better customer collaboration}  \\

        \bottomrule
    \end{tabular}
    \caption{Summarizing why large organizations initiate an agile transformation (RQ1).
             The table presents a hierarchial view of reasons for the organization to change.}
    \label{table:reasonstochange}
\end{table}

The reasons to change were are grouped in three top-level categories: external
issues, internal issues, and communication issues. External issues are related
to the business area which the organization is operating in. Internal issues are
related to the way the organization is managed and software is being developed.
Communication and collaboration issues are related to internal and external
collaboration. Change was motivated by known problems in these categories, such
as known process dysfunctions or communication problems. There was also internal
and external demand for improvement, for instance in speed of delivery and
quality.

\subsubsection{External issues}

The most prominent reason for introducing agile methods was the demand to
deliver software faster and more frequently. A common statement was that it was
necessary to reduce the Time To Market (TTM) [P17, P18, P32, P39, P47]. As
stated in one study: ``The market was moving faster than the development group
was able to deliver'' [P38]. Long cycle time, i.e. time from start of project to
delivery, was seen as a problem in itself [P25, P36]. Some studies indicated
that the organizations set as one of the transformation goals to deliver
software faster [P2, P8]. Other cases described that development was generally
considered as slow [P36, P46, P49].

Commercial competition and market situation was seen as a trigger for
transformation. Especially speed of delivery was seen as a means to respond to
the competition [P9, P38]. As Greening describes: ``The average release duration
peaked at 41 and 35 months, an alarming state that could enable competitors to
gain market share.'' [P19]. In many cases the transformation was seen as
essential for the company to remain competitive [P10, P11, P18, P42, P49]. For
instance, British Telecom realized that ``From a strategic view, there is little
alternative than to change the old ways of working'' [P32].

Demands for change and improvement were also expressed by internal and external
customers. Again, the requests of the customers reflected the need of faster
development and release cycles [P5, P39]. As described in one study: ``Speed of
delivery is of the essence – customers are no longer prepared to wait for 18
months or more for solutions to be delivered'' [P23]. In some cases business
partners expressed mistrust in the original way of conducting development [P40,
P46].

Yet another frequent external factor driving for improvement was the need to
accomodate change during development. Organizations saw agile as a opportunity
to improve responsiveness to change in general [P25, P26, P42].
Some organizations faced difficulties when changes to plans were needed. For
instance, the organizational struture made adapting to changes difficult [52],
and making changes created management overhead [P51]. Other organizations saw an
business advantage if priorities could be changed as required [P8, P39].
Customers and business representatives were in many cases attributed as the
sources of change requests, and the responsiveness to customer requests needed
to be improved [P18, P33, P36, P49].

\subsubsection{Internal issues}

The internal issues relate to how software development is managed in the
organization. In many organizations the project management practice was in a
state of dysfunction, and it needed to be corrected. Management problems were
for instance constantly recurring crisis events [P30], managers ignoring
prioritization of tasks [P38], and reluctance towards focusing on delivering
business value [P49]. Some cases stated that projects simply failed [P21]. Some
organizations wanted to improve generic project management practice, such as
risk management [P8], management of offshoring [P38], and improving the quality
of life of team members [P44]. Several organizations were growing and
experienced that the old development method did not scale up [P16, P30, P50, P51].
There was also a desire to improve team motivation and commitment, and
collaboration between management and developers [P39].

@@

A common problem that was acknowledged in organizations was overhead in process,
which showed as bureucracy and needless costs. Examples on overhead in process
were unproductive meetings [P38], sign-offs and gates [P9], excessive focus on
conformance to process [P40], and change management overhead [P51]. Creating
excessive documentation instead of business value was seen as a problem in some
cases [P24, P35].
From business representatives point of view the processes were unreasonably
costly, while at the same time insufficiently capable of delivering business
value [P36, P46, P49]. In one case business realized that customers did not
request the heavy CMMI based process [P24].
A heavy process resulted often in ignoring the process, only paying ``lip
service'' to the process, or applying its steps retroactively [P6, P9, P49].
Seffernick described how applying the waterfall process had teams ``still
developing the same old way but slower because of the additional process''
[P46].

@@

Finally, the original process model was seen as a cause for quality problems.
Several studies mentioned a demand for better quality, and quality improvement
was set as a goal for the agile transformation [P25, P31, P41, P42]. Improvement
in external quality and customer satisfaction were also anticipated from agile
development [P2, P4]. Problems with technical debt were attributed to the
original process. Some cases reported on prioritizing new feature development
over bugfixes or fixing only the easiest issues [P30, P38].

\subsubsection{Communication and collaboration issues}

@@

Other communication issues were related to requirements handling and difficulty
to focus on the most important aspects of a delivery.
Examples on problems with requirements was too much focus on the documenting
aspect instead of providing understanding [P24, P49], and ignoring the fact that
requirements and understanding are evolving during a project [P4].
In some cases the organization of work and management was such that it hindered
focusing the development [P7], and prioritization of what to work on was not
done efficiently [P38, P39].

@@

A few organizations were experiencing rise of agile methods at grass roots
level, but the agile teams could not collaborate efficiently with other parts of
a waterfall organization. Tensions arose when teams using different development
approaches attempted to work together on products [P10]. On a team level agile
had enabled some flexibility, but that flexibility could not be leveraged on the
scale of the entire organization [P26]. This kind of experiences drove the
organization to deploy agile on a full scale.

@@

In a few cases the organization initailly lacked strict process. For instance
Long describes the initial state in the following way: ``Project scope was in
constant flux. Well-meaning market and product owners routinely inserted new
requirements during product development.'' [P30].

Some authors describe the developement being originally organized in silos. For
instance Ranganath describes that ``Teams operated in silos. Collaboration,
shared purpose and ownership for outcomes were hard to find'' [P40].

>>>>

As stated by Research Question 1 we investigated the reasons organizations start
a transformation process towards agile software development. Most primary
studies mentioned reasons for starting the transformation, and only 7 cases of
42 did not mention any reasons. In the cases giving no specific reason to change
it was simply mentioned that management had decided to change the organization.
The reported reasons for starting an organizational change towards agile are
summarized in figure \ref{fig:reasonstochange_summary}.

@@

The most prominent reason to start the transformation was that business demanded
shorter lead times and faster delivery. Other top reasons to start change were
dysfunction in management, overhead in the current process, problems in
scheduling, and demand to improve quality.

\begin{figure}[h]
  \begin{center}
    \includegraphics{graphics/reasonstochange_summary.pdf}
    \caption{Reasons why organizations started an agile transformation}
    \label{fig:reasonstochange_summary}
  \end{center}
\end{figure}

\subsubsection{Business demands speed and flexibility}

Most cases reported that a major reason to change was that business demanded
improvement. Most companies wanted to improve their competitiveness, or even
feared that they are losing competition. The way to respond was to improve
delivery speed and responsiveness to change. Table
\ref{table:reasonstochange_business} presents business related reasons to
introduce an agile model.

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Business reasons for starting a change   &  Described in  \\
        \midrule
        Time to market must be reduced    &  P17, P18, P32, P38, P39, P47  \\
        Delivery is considered slow       &  P2, P8, P25, P36, P46, P49  \\
        Something must be done to remain competitive  &
                P9, P10, P11, P18, P19, P32, P38, P42, P49  \\
        Business partners demand faster delivery  &  P5, P23, P39, P40, P46  \\
        Need to accomodate change during projects  &
                P4, P8, P18, P25, P26, P33, P36, P39, P42, P49, P51, P52  \\
        \bottomrule
    \end{tabular}
    \caption{Business related reasons for starting an agile transformation}
    \label{table:reasonstochange_business}
\end{table}

The most frequently mentioned reason for introducing agile methods was the demand to
deliver software faster and more frequently. A common statement was that it was
necessary to reduce the time to market (TTM) [P17, P18, P32, P39, P47]. As
stated in one study: ``The market was moving faster than the development group
was able to deliver'' [P38]. Long cycle time, i.e. time from start of project to
delivery, was seen as a problem in itself [P25, P36]. Some studies indicated
that the organizations set as one of the transformation goals to deliver
software faster [P2, P8]. Other cases described that development was generally
considered as slow [P36, P46, P49].

Commercial competition and market situation was seen as a trigger for
transformation. Especially speed of delivery was seen as a means to respond to
the competition [P9, P38]. As Greening [P19] describes: ``The average release
duration peaked at 41 and 35 months, an alarming state that could enable
competitors to gain market share.''. In many cases the transformation was seen
as essential for the company to remain competitive [P10, P11, P18, P42, P49].
For instance, British Telecom realized that ``From a strategic view, there is
little alternative than to change the old ways of working'' [P32].

Demands for change and improvement were also expressed by internal and external
customers. Again, the requests of the customers reflected the need of faster
development and release cycles [P5, P39]. As described in one study: ``Speed of
delivery is of the essence -- customers are no longer prepared to wait for 18
months or more for solutions to be delivered'' [P23]. In some cases business
partners expressed mistrust in the original way of conducting development [P40,
P46].

Yet another factor pushing for improvement was the need to accomodate change
during development. Organizations saw agile as an opportunity to improve
responsiveness to change in general [P25, P26, P42].
Some organizations faced difficulties when changes to plans were needed. For
instance, the organizational struture made adapting to changes difficult [52],
and making changes created management overhead [P51]. Other organizations saw an
business advantage in enabling priorities to be changed as required [P8, P39].
Domain knowledge will deepen during projects and environments will inevitably
change, and therefore the requirements should not stay fixed [P4].
Customers and business representatives were in many cases attributed as the
sources of change requests, and the responsiveness to customer requests needed
to be improved [P18, P33, P36, P49].

\subsubsection{Management issues}

The way software development was managed was typically seen as problematic. The
problems related to project management, people management, and managing
schedules. Management related problems that the transformation was hoped to
correct are listed in table \ref{table:reasonstochange_management}.

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Management problems           &  Described in  \\
        \midrule
        Software development is managed badly   &  P8, P21, P30, P38, P44, P49  \\
        Command and control style management    &  P9, P25, P30, P40, P44  \\
        Scheduling and estimation is difficult  &  P16, P19, P21, P30, P33, P40, P41, P51  \\
        \bottomrule
    \end{tabular}
    \caption{Management problems stated as reasons for organizational change}
    \label{table:reasonstochange_management}
\end{table}

In many organizations the project management practice was in a state of
dysfunction, and it needed to be corrected. Project management problems included
constantly recurring crisis events [P30], managers ignoring prioritization of
tasks [P38], and reluctance towards focusing on delivering business value [P49].
Some cases stated that projects simply failed [P21]. Some organizations wanted
to improve generic project management practice, such as risk management [P8],
management of offshoring [P38], and generally improving the quality of life of
team members [P44].

@@

\subsubsection{Process problems}

The old way of working was acknowledged to have problems which needed to be
corrected. A summary of weaknesses in the original process is presented in table
\ref{table:reasonstochange_process}.

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.51\textwidth}
                     >{\raggedright\arraybackslash}p{0.41\textwidth} }
        Problems in the old process           &  Described in  \\
        \midrule
        Excess and overhead in process     &  P9, P24, P35, P36, P38, P40, P46, P49, P51  \\
        The process is being ignored       &  P6, P9, P46, P49  \\
        Late integration, testing, and feedback  &  P5, P25, P33, P40, P50  \\
        New practices can not be integrated in the old process  &  P10, P26  \\
        The old process does not scale up  &  P16, P30, P50, P51  \\
        \bottomrule
    \end{tabular}
    \caption{Problems because of which the old process should be changed}
    \label{table:reasonstochange_process}
\end{table}

A most typical known problem was overhead in process, which showed as bureucracy and
needless costs. Examples on overhead in process were unproductive meetings
[P38], sign-offs and gates [P9], excessive focus on conformance to process
[P40], and change management overhead [P51]. Creating excessive documentation
instead of business value was seen as a problem in some cases [P24, P35]. From
business representatives point of view the processes were unreasonably costly,
while at the same time insufficiently capable of delivering business value [P36,
P46, P49]. In one case business realized that customers did not see any benefit
in the heavy CMMI based process [P24].

A heavy process resulted often in ignoring the process, only paying ``lip
service'' to the process, or applying its steps retroactively [P6, P9, P49].
Seffernick [P46] describes how applying the waterfall process had teams ``still
developing the same old way but slower because of the additional process''.

@@

A few organizations were experiencing rise of agile methods at grass roots
level, but the agile teams could not integrate efficiently with other parts of
a waterfall organization. Tensions arose when teams using different development
approaches attempted to work together on products [P10]. On a team level agile
had enabled some flexibility, but that flexibility could not be leveraged on the
scale of the entire organization [P26].
This kind of experiences drove the organization to deploy agile on a full scale.

Still one problem was that some growing organizations experienced that the old
development method did not scale up [P16, P30, P50, P51].

\subsubsection{Other reasons to change}

There were still a number of problems that were given as reasons to change.
In addition to the above mentioned problems, organizations were striving for
better collaboration and quality. Table \ref{table:reasonstochange_other} lists
other issues that drove organizations to change.

\begin{table}[b]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.61\textwidth}
                     >{\raggedright\arraybackslash}p{0.31\textwidth} }
        Reasons to change           &  Described in  \\
        \midrule
        Teams can not collaborate properly   &  P21, P38  \\
        Silos and internal boundaries in organization   &  P27, P40, P46  \\
        Lack of customer collaboration       &  P4, P26, P40  \\
        Demand for better quality            &  P2, P4, P25, P31, P30, P38, P41, P42  \\
        \bottomrule
    \end{tabular}
    \caption{Other reasons why organizations wanted to change the way of working}
    \label{table:reasonstochange_other}
\end{table}

@@

@@

% There was also a desire to improve team motivation and commitment, and
% collaboration between management and developers [P39].

% Examples of problems with requirements was too much focus on the documenting
% aspect instead of providing understanding [P24, P49], and ignoring the fact that
% requirements and understanding are evolving during a project [P4].
% 
% In some cases the organization of work and management was such that it caused
% lack of focus the development [P7], and prioritization of what to work on was
% not done efficiently [P38, P39].

Lastly, the original model was seen as a cause for quality problems.
Several studies mentioned a demand for better quality, and quality improvement
was set as a goal for the agile transformation [P25, P31, P41, P42]. New feature
development was being prioritized over bugfixes, and technical debt was
developing [P30, P38]. Improvement in external quality and customer satisfaction
were anticipated from agile development [P2, P4].

@@

In a few cases the organization initailly lacked strict process. For instance
Long [P30] describes the initial state in the following way: ``Project scope was
in constant flux. Well-meaning market and product owners routinely inserted new
requirements during product development.''.

Some authors describe the developement being originally organized in silos. For
instance, Ranganath [P40] describes that ``Teams operated in silos.
Collaboration, shared purpose and ownership for outcomes were hard to find''.
